Extended Data

### 2021-12-03 doc/manuscript/Dow_manuscript_Nature.Rmd
working on code/data statements
#93

<!--Climate -->
For the dendrometer band and leaf phenology analyses, climate data corresponding to the measurement periods were obtained from local weather stations at each focal site. 
For SCBI, weather data were obtained from a meteorological tower adjacent to the ForestGEO plot, via the ForestGEO Climate Data Portal v1.0 (https://forestgeo.github.io/Climate/)[@anderson-teixeira_forestgeo_2020]. 
The R package *climpact* (see www.climpact-sci.org)[@benestad_empiricalstatistical_2008] was used to plot temperatures for visual inspection and to identify readings that were >3 standard deviations away from yearly means, which were labeled as outliers and removed from the dataset. 
Gaps in the SCBI meteorological tower data were subsequently filled using temperature readings obtained from a National Center for Environmental Information (NCEI) weather station located in Front Royal, Virginia (https://www.ncdc.noaa.gov/cdo-web/datasets/GHCND/stations/GHCND:USC00443229/detail). 
Daily temperature records for Harvard Forest, which had already been gap-filled based on other local records, were obtained from the Harvard Forest weather station[@boose_shaler_2021; @boose_fisher_2021]. 
For each site, we used records of daily maximum ($T_{max}$) and minimum temperatures ($T_{min}$).

@@

# Acknowledgements {-#Acknowledgements} 
We gratefully acknowledge all researchers who assisted with data collection in the field and laboratory, particularly Tsun Fung Au, Joshua Bregy, James Dickens, Karen Heeter, Anna Hennage, Daniel King, James McGee, Benjamin Lockwood, Jennifer McGarvey, Victoria Meakem, Josh Oliver, Jessica Shue, Karly Schmidt-Simard, Brandon Strange, Alyssa Terrell, Brynn Taylor, Michael Thornton, Senna Robeson, Matt Wenzel, and Luke Wylie. 
Thanks to David A. Orwig and members of the ForestGEO Ecosystesms & Climate Lab at SCBI for helpful feedback.
The research was funded by ForestGEO (Smithsonian). 
Collection of tree-ring samples was funded by a USDA Agriculture and Food Research Initiative grant 2017-67013-26191 and from the Indiana University Vice Provost for Research Faculty Research Program.

@@

# Data Availability Statement {-#Data}

# Code Availability Statement {-#Code}
Data were analysed in the open source statistical software R (version 4.0).
We used packages *Rdendrom* (https://github.com/seanmcm/RDendrom/), *climwin* v.1.2.3 (https://cran.r-project.org/web/packages/climwin/index.html), *climpact* (see www.climpact-sci.org)[@benestad_empiricalstatistical_2008], *dplR* v.1.0.2, and *bootRes* v1.2.4. 
All custom code is available through the EcoClimlab GitHub repository (https://github.com/EcoClimLab/growth_phenology) and archived in Zenodo (DOI: [TBD]).

>>>>

<!--Weather data -->
For the dendrometer band and leaf phenology analyses, climate data corresponding to the measurement periods were obtained from local weather stations at each focal site. 
For SCBI, weather data were obtained from a meteorological tower adjacent to the ForestGEO plot, via the ForestGEO Climate Data Portal v1.0 (https://forestgeo.github.io/Climate/)[@anderson-teixeira_forestgeo_2020]. 
The R package *climpact* (see www.climpact-sci.org)[@benestad_empiricalstatistical_2008] was used to plot temperatures for visual inspection and to identify readings that were >3 standard deviations away from yearly means, which were labeled as outliers and removed from the dataset. 
Gaps in the SCBI meteorological tower data were subsequently filled using temperature readings obtained from a National Center for Environmental Information (NCEI) weather station located in Front Royal, Virginia (https://www.ncdc.noaa.gov/cdo-web/datasets/GHCND/stations/GHCND:USC00443229/detail). 
Daily temperature records for Harvard Forest, which had already been gap-filled based on other local records, were obtained from the Harvard Forest weather station[@boose_shaler_2021; @boose_fisher_2021]. 
For each site, we used records of daily maximum ($T_{max}$) and minimum temperatures ($T_{min}$).

@@

# Acknowledgements {-#Acknowledgements} 
We gratefully acknowledge all researchers who assisted with data collection in the field and laboratory, particularly Tsun Fung Au, Joshua Bregy, James Dickens, Karen Heeter, Anna Hennage, Daniel King, James McGee, Benjamin Lockwood, Jennifer McGarvey, Victoria Meakem, Josh Oliver, Jessica Shue, Karly Schmidt-Simard, Brandon Strange, Alyssa Terrell, Brynn Taylor, Michael Thornton, Senna Robeson, Matt Wenzel, and Luke Wylie. 
Thanks to David A. Orwig and members of the ForestGEO Ecosystems & Climate Lab at SCBI for helpful feedback.
The research was funded by ForestGEO (Smithsonian). 
Collection of tree-ring samples was funded by a USDA Agriculture and Food Research Initiative grant 2017-67013-26191 and from the Indiana University Vice Provost for Research Faculty Research Program.

@@

# Data Availability Statement {-#Data} 
Weather data for SCBI were obtained from the ForestGEO Climate Data Portal v1.0 (https://github.com/forestgeo/Climate/tree/master/Climate_Data/Met_Stations/SCBI), which is archived in Zenodo (DOI: 10.5281/zenodo.3958215), and the National Center for Environmental Information (NCEI) weather station located in Front Royal, Virginia (https://www.ncdc.noaa.gov/cdo-web/datasets/GHCND/stations/GHCND:USC00443229/detail), downloaded on [DATE].
Weather data for Harvard Forest are available through the Harvard Forest Data Archive (https://harvardforest.fas.harvard.edu/harvard-forest-data-archive) <!--specify link, version-->.
Climate data were obtained from CRU v.4.04 via the ForestGEO Climate Data Portal v1.0 (https://github.com/forestgeo/Climate/tree/master/Climate_Data/CRU), which is archived in Zenodo (DOI: 10.5281/zenodo.3958215).

# Code Availability Statement {-#Code}
Data were analysed in the open source statistical software R (version 4.0).
We used packages *Rdendrom* (https://github.com/seanmcm/RDendrom/), *climwin* v.1.2.3 (https://cran.r-project.org/web/packages/climwin/index.html), *climpact* (see www.climpact-sci.org)[@benestad_empiricalstatistical_2008], *dplR* v.1.0.2, and *bootRes* v1.2.4. 
All custom code is available through the EcoClimlab GitHub repository (https://github.com/EcoClimLab/growth_phenology) and archived in Zenodo (DOI: *[TBD]*).
